### 2020-12-04 paper/localpsf.tex
did many good edits to the intro

In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
\begin{equation}
\label{eq:kernel_representation}
(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$. We focus on operators of this form which have the following properties:
\begin{description}
\item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$. Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. 
\item[Non-negative:] $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
\item[Local support:] $A(y, x)$ is supported near $y=x$.
\item[Local translation invariance:] $A(y+h, x+h) \approx A(y,x)$
if $h$ is not too large.
\end{description}

If $\mathcal{A}$ were translation invariant (that is, if $A(y+h, x+h) = A(y,x)$ for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
\begin{equation}
\label{eq:convolution_kernel}
	\varphi_x(y) := A(y+x,x).
\end{equation}
The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$. Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes. Locally translation invariant operators may therefore be approximated by spatially varying weighted sums of convolution operators of the form
\begin{equation}
\label{eq:product_convolution}
	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
\end{equation}
here $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
\begin{itemize}
	\item $\varphi_i := \varphi_{x_i}$ is given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
	\item The functions $w_i(x)$ are spatially varying weighting function that interpolate the convolution kernels $\varphi_i$.
\end{itemize}
Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution. The more locally translation invariant an operator is, the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate approximation.

Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by
\begin{equation}
\label{eq:kernel_entries}
	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a hierarchical matrix (H-matrix) representation of $\widetilde{\mathcal{A}}$. Using this H-matrix representation, one can invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations with $\widetilde{\mathcal{A}}$, using fast H-matrix arithmetic.

For background on product-convolution approximations, we recommend reading the following papers: [CITES]. For background on H-matrices, we recommend reading the following papers: [CITES].

@@

@@

\section{Poisson PSF interpolation}

>>>>

We present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
\begin{equation}
\label{eq:kernel_representation}
(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$. We focus on operators of this form which have the following properties:
\item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$. Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors. That is, evaluation of the functions
	u \mapsto\mathcal{A}u \quad \text{and} \quad v \mapsto\mathcal{A}^Tv.
\item[Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
	A(y,x) \ge 0.
\item[Local support:] $A(y, x)$ is supported near $y=x$.
\item[Local translation invariance:] If $h$ is not too large, then
\begin{equation}
	\label{eq:local_translation_invariance}
	A(y+h, x+h) \approx A(y,x).
\end{equation}
Local translation invariance is illustrated in Figure FIG. 
%We qualify the degree of local translation invariance with the words ``perfectly'', ``highly'', and ``poorly''.
%\begin{itemize}
%\item If equality holds in \eqref{eq:local_translation_invariance} for all $x$, $y$, and $h$, then we say $\mathcal{A}$ is \emph{perfectly translation invariant}. 
%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is small for a large set of points $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{highly locally translation invariant}. 
%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is large, or if the discrepancy is only small for a small set of $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{poorly locally translation invariant}.
%\end{itemize} 
If $\mathcal{A}$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
\begin{equation}
\label{eq:convolution_kernel}
	\varphi_x(y) := A(y+x,x).
\end{equation}
The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source at some fixed point $x \in \Omega$, then translating the resulting impulse response function re-center it at zero instead of $x$. Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes. Locally translation invariant operators may therefore be approximated by spatially varying weighted sums of convolution operators of the form
\begin{equation}
\label{eq:product_convolution}
	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
\end{equation}
where $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
\begin{itemize}
	\item $\varphi_i := \varphi_{x_i}$ is given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
	\item The functions $w_i(x)$ are spatially varying weighting function that interpolate the convolution kernels $\varphi_i$.
\end{itemize}
Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution. The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.

The core content of this paper is specifying (a) how we choose the points, $x_i$, and compute the convolution kernels, $\varphi_i$ (Section \ref{sec:choosing_sample_points}), and (b) what weighting functions $w_i$ that we use (Section \ref{sec:weighting_functions}). Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
\begin{equation*}
	\eta = \mathcal{A} \left(\sum_{i=1}^m \delta_{x_i}\right).
\end{equation*}
The batch of points $x_i$ are chosen so that the support of $\mathcal{A}\delta_{x_i}$ and the support of $\mathcal{A}\delta_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$. Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one vector, we recover many $\varphi_i$. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached. The supports of the impulse responses $\mathcal{A}\delta_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions. This procedure for estimating the support of $\mathcal{A}\delta_x$ a-priori is the only part of the paper that requires non-negativity of the integral kernel. The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.

%A small number of terms in \eqref{eq:product_convolution} are required to accurately approximate a highly locally translation invariant operator, while a large number of terms are required to accurately approximate a poorly locally translation invariant operator.

Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
\begin{equation}
\label{eq:kernel_entries}
	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a hierarchical matrix (H-matrix) representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method [CITE]. Once the H-matrix representation is formed, one can invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations with $\widetilde{\mathcal{A}}$, using fast H-matrix arithmetic.

For background on product-convolution approximations, we recommend reading the following papers: [CITES]. In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}. In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid. However,

@@

% Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.

%Specifically, since the integral kernel for $\mathcal{A}$ is non-negative, $\langle\mathcal{A},\delta_x\rangle^*$ may be interpreted as a scaled probability distribution. We compute the mean and covariance of this probability distribution for all points $x \in \Omega$ simultaneously, by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions. We then estimate the support of $\langle\mathcal{A},\delta_x\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution.

%We interpret $\langle\mathcal{A},\delta_{x_i}\rangle^*$ as a scaled probability distribution, and estimate the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution (this is why we require a non-negative kernel).  The mean and covariance of $\langle\mathcal{A},\delta_{x}\rangle^*$ can be simultaneously computed for all points $x \in \Omega$ by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.

The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$. Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.

@@

\section{Poisson PSF interpolation}
\label{sec:weighting_functions}
